---
title: "Final Project"
author: "Qianhui Ni and Wani Qiu"
date: "4/7/2022"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
library(here)
library(magrittr) # for the `%>%` operator
library(rstan)
rstan_options(auto_write = TRUE) # save compiled STAN object
library(posterior)
library(bayesplot)
library(brms)  # simplify fitting Stan GLM models
library(modelsummary)  # table for brms
theme_set(theme_classic() +
theme(panel.grid.major.y = element_line(color = "grey92")))
```


## Research Question

Do children differentiate the complexity of the information they teach based on the maturity of the audience?

## Description of each variable in our preliminary analysis

Age: five-years-old and seven-years-old

Response: basic or complex information

Trials: number of trials per participant


##  Model and the priors  

### Model:  

$$
  \begin{aligned}
    z_j^A & \sim \mathrm{Bin}(N_j^A, \theta_j^A) \\
    z_j^B & \sim \mathrm{Bin}(N_j^B, \theta_j^B) \\
    \theta_j^A & \sim \mathrm{Beta2}(\mu, \kappa) \\
    \theta_j^B & =\Phi(q_j) \\
    q_j & = \Phi^{-1}(\theta_j^A) + \delta_j \\
    \delta_j & \sim N(\mu_{\delta}, \tau_{\delta})
  \end{aligned}
$$

### Prior:

$$
\begin{aligned}
 \mu & \sim \mathrm{Beta}(2,2) \\
 \kappa & \sim \mathrm{Gamma}(0.01,0.01)\\
 \mu_{\delta} & \sim N(0,1) \\
 \tau_{\delta} & \sim N(0,1)
\end{aligned}
$$
$j$:subject (subject 1, 2, ..., 68)  

$\theta_j$: individual subject's prob of transmitting complex info

$N_j$:the number of trails


```{r echo=FALSE}
#Import data
alldata <- read.csv('Bayes_Project_data.csv')

five <- subset(alldata, Age.Group==5)
seven <- subset(alldata, Age.Group==7)

```

# 5-year-olds

```{r include=FALSE}
five_ntrial_adu <- five$Atrial
five_Acomplex <- five$Acomplex
five_ntrial_baby <- five$Btrial
five_Bcomplex <- five$Bcomplex
fit1 <- stan(
  file = here("hierarchical_bern_within_study.stan"),
  data = list(J = 34, yc = five_Acomplex, Nc = five_ntrial_adu,
  yt = five_Bcomplex, Nt = five_ntrial_baby),
  seed = 1234, # for reproducibility
  iter = 4000,
  cores = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 12)
)
```
## Convergence Check for 5yos

```{r echo=FALSE, fig.height=10}
mcmc_intervals(fit1,
               regex_pars = "theta")
```

### Trace plots of key parameters
```{r echo=FALSE}

mcmc_trace(fit1, pars = c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```


### Rank histograms of key parameters
```{r echo=FALSE}
mcmc_rank_hist(fit1, pars = c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```


### Posterior distribution of key model parameter mu:
```{r echo=FALSE}
mcmc_dens(fit1, pars=c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```

### Table of the posterior distributions of key parameters

```{r echo=FALSE}
fit1%>%
as_draws() %>%
  summarize_draws() %>%
  knitr::kable(digits=2)
```

From the trace plots and the rank histograms, it seems like there is good mixing of the 4 chains as they frequently cross each other. The histograms for the four chains have similar average ranks and are roughly uniform, which also suggests good mixing. The ess_bulk for all of the key parameters is larger than 1000, and the rhat value for all key parameters are less than 1.01.  

## 7-year-olds
```{r message=FALSE, warning=FALSE, include=FALSE}
seven_ntrial_adu <- seven$Atrial
seven_Acomplex <- seven$Acomplex
seven_ntrial_baby <- seven$Btrial
seven_Bcomplex <- seven$Bcomplex
fit2 <- stan(
  file = here("hierarchical_bern_within_study.stan"),
  data = list(J = 34, yc = seven_Acomplex, Nc = seven_ntrial_adu,
  yt = seven_Bcomplex, Nt = seven_ntrial_baby),
  seed = 1234, # for reproducibility
  iter = 4000,
  cores = 4,
  control = list(adapt_delta = 0.99, max_treedepth = 12)
)

fit2

```


### Convergence Check for 7yos

```{r echo=FALSE, fig.height=10}
mcmc_intervals(fit2,
               regex_pars = "theta")
```


### Trace plots of key parameters
```{r echo=FALSE}
#trace plot of key parameter mu
mcmc_trace(fit2, pars = c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```


### Rank histograms of key parameters
```{r echo=FALSE}
mcmc_rank_hist(fit2, pars = c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```


### Posterior distribution of key model parameter mu:
```{r echo=FALSE}
mcmc_dens(fit2, pars=c("mu", "mu_adult_basic", "mu_baby_complex", "mu_baby_basic"))
```

### Table of the posterior distributions of key parameters

```{r echo=FALSE}
fit2%>%
as_draws() %>%
  summarize_draws() %>%
  knitr::kable(digits=2)
```

From the trace plots and the rank histograms, it seems like there is good mixing of the 4 chains as they frequently cross each other. The histograms for the four chains have similar average ranks and are roughly uniform, which also suggests good mixing. The ess_bulk for all of the key parameters is larger than 1000, and the rhat value for all key parameters are less than 1.01.  

## Interpretation of Results

### Within-Group Comparisons

#### 5-year-olds

When teaching an adult, 5-year-olds were equally likely to teach basic and complex information. The estimated difference in probability is -0.09, 90% CI[-0.23, 0.04]. When teaching a baby, 5-year-olds were also equally likely to teach basic and complex information. The estimated difference in probability is -0.16, 90% CI[-0.32, 0.01]. 

The estimated probability for 5-year-olds teaching complex information to the adult is 0.45, 90% CI [0.38, 0.52]. Since this interval includes 0.5, it indicates that they are teaching complex information to the adult at chance. The estimated probability for 5-year-olds teaching basic information to the baby is 0.58, 90% CI [0.49, 0.66]. Since this interval includes 0.5, it indicates that they are also teaching basic information to the baby at chance. 

#### 7-year-olds
When teaching an adult, 7-year-olds were more likely to teach complex information than basic information. The estimated difference in probability is 0.23, 90% CI[0.08, 0.37]. When teaching a baby, 7-year-olds were equally likely to teach basic and complex information. The estimated difference in probability is -0.13, 90% CI[-0.39, 0.13]. 

The estimated probability for 7-year-olds teaching complex information to the adult is 0.61, 90% CI [0.54, 0.69]. Since this interval does not include 0.5, it indicates that they are teaching complex information to the adult above chance. The estimated probability for 7-year-olds teaching basic information to the baby is 0.57, 90% CI [0.43, 0.69]. Since this interval includes 0.5, it indicates that they are teaching basic information to the baby at chance. 

#### Age Group Comparisons

From the posterior distributions, it seems like 7-year-olds are better than 5-year-olds at teaching complex information to the adult. The mean amount of complex information taught to adults for 7-year-olds is 0.61, 90% CI [0.54, 0.69], whereas the mean amount of complex information taught to adults for 5-year-olds is 0.45, 90% CI[0.38, 0.52]. However, there does not seem to be an age difference in the probability of basic information taught to the baby audience. The estimated probability of 7-year-olds teaching basic information to the baby is 0.57, 90% CI [0.43, 0.69]. Similarly, the estimated probability of 5-year-olds teaching basic information to the baby is 0.58, 90% CI [0.49. 0.66].

This indicates that compared to 5-year-olds, 7-year-olds are significantly better at teaching complex information to the adult. On the other hand, 7-year-olds and 5-year-olds were similar in their probability of teaching basic information to the baby.